%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% chapter01 - Cyber-physical Systems
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
% >>>>>>>>>>>>>>> PLEASE NOTE <<<<<<<<<<<<<<<
%
% This file is not stand-alone compileable as it is, to make it compileable while writing uncomment the preamble below.
% In this case, you also have to uncomment the begin/end document statements.
% You can outcomment the preamble and the begin/end document statements again or erase them when handing in your contribution.
%
% If you use BibTex for your bibliography, please use \putbib[bibliography] to print your reference (see end of this file).
%
% you can use paths relative to your chapter dir, e.g. \figure{assets/fig1}.
%
% >>>>>>>>>>>>>>>>>>>><<<<<<<<<<<<<<<<<<<<<<<

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% you can uncomment the following preamble during development to make this file compileable.
%% Note that you need the svmult.cs file inside your chapter root dir to make this work.
%% Also note that if you need additional packages etc., you can add them here, but please
%% mark them somehow so the editor of this book knows you need them in the final book.
%% When you hand in your contribution, please uncomment or remove the preamble again.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% start of preamble
\documentclass[
graybox,
envcountchap,
natbib
]{svmult}

\usepackage[utf8]{inputenc}
\usepackage{type1cm}        % activate if the above 3 fonts are not available on your system

\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

\usepackage{newtxtext}       % 
\usepackage{newtxmath}       % selects Times Roman as basic font

\usepackage{natbib}
\usepackage{footmisc}

% Additional packages added. Add necessary packages here.
\usepackage[english]{babel}
\usepackage{siunitx}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{url}
\usepackage{mathtools}
\usepackage{lipsum}
\usepackage{import}
\usepackage{bibunits}
\usepackage{acronym}
\usepackage[nottoc]{tocbibind}
\usepackage{numberpt}

\newcommand*{\CHAPTERSROOT}{../.}	% root path for chapters.
\newcommand*{\chapterprefix}{01}	% your chapter number.

\makeindex % used for the subject index
%%end of preamble

%uncomment the \begin{document} statement to make this file stand-alone compileable.
\begin{document}

\begin{bibunit}
	
	\title*{Cyber-physical Systems}
	\author{Carlos Varela and Damien Zufferey}
	
	\institute{
		Carlos Varela \at Rensselaer Polytechnic Institute, USA, \email{cvarela@cs.rpi.edu}
		\and Damien Zufferey \at Max Planck Institute for Software Systems, Germany \email{zufferey@mpi-sws.org}
		\and Gul Agha \at University of Illinois, USA, \email{agha@illinois.edu}
		\and Takuo Watanabe \at Tokyo Institute of Technology, Japan, \email{takuo@c.titech.ac.jp}
		\and Shoji Yuen \at Nagoya University, Japan, \email{yuen@is.nagoya-u.ac.jp}
		\and Marjan Sirjani \at Mälardalen University, Sweden, \email{marjan.sirjani@mdh.se}
	}
	\maketitle
	
	\abstract{Please place your abstract here.}
	
	%% content
	\section{Programming Cyber Physical Systems: Introduction}\label{sec:1} %(3pp)

         CPS = communication, computation, and control

        \section{Models} %(3pp)

        \subsection{Control theory}

         Control theory approach (plant model and feedback controller)

         \subsection{Hybrid/Timed Automata}

         \subsection{Process Algebra}

         \subsection{Dynamic Logic}

         \subsection{Markov Processes}

         \section{Communication} %(4pp)

         \subsection{Synchronous rendez-vous communication}

         fits well with automata models, good abstraction for slow dynamics and fast

         \subsection{Actor Model}

         \subsection{Functional Reactive Programming}

         communication using streams

         \subsection{Session Types}

         % cvarela:  really needed here?  or just a ref to chapter 2?

         \subsection{Networked Control System}

         centralized vs distributed controller

         \section{Communication Imperfection and Control} %(2pp)

         analogous to consistency question in distributed system but about the freshness of physical informations

         \subsection{Clock synchronization}

         a basis on which other functionality are built

         \subsection{Robustness against perturbation of communication}

         example: work on Skorokhod Distance by Ichiro Hasuo

         \subsection{Efficient use of communication}

         example: work by Sebastian Trimpe on event triggered and wireless control systems

         \subsection{Interpolation and extrapolation}

         example: tfs2 ROS library use timestamped messages to interpolate/extrapolate data

         \section{Impact of the Physical Layer and the Communication} %(2pp)

         \subsection{Point-to-point vs broadcast communication}

         \subsection{Large number of processes}

         swarm / sensor networks

         \subsection{Line-of-sight / limited range}

         gossip protocol

         \section{Open Research Questions} %(2-5pp)

         \subsection{Discrete vs continuous systems}

         Process algebra, actors, and others model discrete systems.
         Differential equations, calculus, and others model the physical world.

         \begin{itemize}
         \item How to incorporate continuous time and space into concurrency models?
         \item How to go beyond bi-simulation and observational equivalence to reason about continuous systems?
         \end{itemize}
         
         \subsection{Hardware/Software interface}

         Cyber-physical systems have a hardware (physical) and a software (cyber) component
        
         \begin{itemize}
         \item How to model SW/HW interface?
         \item Are real-time properties properly modeled?
         \item  How to tackle complexity via abstraction without losing key properties?  (Translucency vs black box approaches)
         \item How to verify infinite-space systems?
         \end{itemize}

         \subsection{Dealing with uncertainty}

         Stochastic nature of cyber-physical systems, e.g., weather, requires probabilistic approach.
        
         \begin{itemize}
         \item Are heterogeneous latencies, failure modes properly accounted for?
         \item How to accurately model, quantify, propagate uncertainty?
         \item Need for statistical reasoning libraries suitable for interactive/automated proof assistants.
         \end{itemize}

         \subsection{Modal logics and reasoning}
        
         \begin{itemize}
         \item  Is first-order logic sufficient to reason about cyber-physical systems?
         \item Are spatial logics, temporal logics, and combinations thereof better suited for specifying and reasoning about CPS?
         \item What are their expressive power?   Are there efficient decision procedures?  (SAT modulo theory approaches?)
         \end{itemize}

         \subsection{Robust control}

         Can we reason about properties of feedback loop control systems incorporating the above (hybrid, SW/HW/Network, uncertainty/failures, modal logics/reasoning)?
         
         \subsection{AI and data-driven systems}

         As more CPS are model-driven and more models are data-driven, how can we trust these systems?  Are there inherent theoretical limits to dynamic data-driven applications and systems? (e.g., Cramer-Rao lower bounds, etc.)

	% \section*{Appendix}\label{appendix}
	
	% Please place your appendix content here, if applicable.
	
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%% For your bibliography, you should use a bibtex .bib file and include it here.
	%% Note that the final reference lists styling might differ because it'll be styled in unified book layout.
	
	% \biblstarthook{
	%	text inserted here will be printed before the actual list of references, but only if there is at least one reference to %display. Delete this section if you don't need it.
	%}
	
	% \nocite{*}		%% uncomment if uncited references should be listed in the bibliography.
	
	%% uncomment and state path to your .bib to use a bibtex file as your bibliography.
	%% NOTE: relative paths don't work in \putbib => During development, you might delete the "\CHAPTERSROOT/chapter\chapterprefix/" part to refer to your bib file. When you're done, please make this path absolute by adding the prefix again.
	%%
	% \putbib[\CHAPTERSROOT/chapter\chapterprefix/bibliography] %
	
\end{bibunit}
	
%% uncomment the \end{document} statement to make this file stand-alone compileable.
\end{document}